Supplementary Material

### 2022-08-09 supplement.tex
Update supplement.tex

\usepackage[letterpaper]{geometry}
\usepackage{authblk}
\usepackage{graphicx}

@@

One important experimental detail that all previous plots took into account was recording the flow profiles while the outlet tubing from the phantom was submerged in solution, as shown in Figs. 2 and 3. This was performed after visualizing the dramatic decrease in flow observed when the outlet line was elevated and the solution dripped into the outlet vial. The release of the drop from the end of the tubing caused a large decrease in pressure, resulting in a flow decrease. This was observed with both flow control systems as shown in Fig.~\ref{fig:supplement_outlet}, meaning that it is an important experimental consideration for any microfluidic setup. These trials included approximately five-minute duration with the outlet line elevated and allowed to drip into the outlet vial, followed by submersion of the outlet line in the solution for another five minutes. Thus, the large spike in relative flow corresponds to the time point where the outlet line was manually submerged in solution.

% Figure S1 - Outlet Line Submersion
\begin{figure}
    \includegraphics[width=\textwidth]{FigureS1.pdf}
    \caption {
        Impact of submerging the microfluidic device outlet tubing for both flow control systems at 2.4 and 4.8 mm/s as measured with the inline flow sensor (blue) and single-exposure LSCI (red). These trials show approximately five minutes of effluent dripping into the collection vial, as evidenced by the rapid periodic flow drops, followed by approximately five minutes of operation with the outlet tubing submerged in solution. All measurements were normalized to the average flow during the submerged segment of each experiment and compared to the ideal flow (black).
    }
    \label{fig:supplement_outlet}
\end{figure}

Some general trends were observed from these trials. First, the magnitude of the flow decrease was larger for the syringe pump versus the pressure system. Because the pressure system has a higher compliance than the syringe pump, high frequency perturbations are dampened more by the pressure system. The frequency of the droplets was also very regular within trials. This suggests that the drop must reach a similar critical volume regardless of flow rate to overcome the surface tension required to release the drop from the tubing. The mean volume of the droplets can be estimated using the absolute flow rate from the flow sensor and the time between drops, and a standard deviation of <0.5 µL in drop volume was observed within each trial. The magnitude of the flow decrease was also larger for the slower flow (2.4 mm/s) versus the faster flow (4.8 mm/s) for all trials. This is likely due to the larger pressure differential generated at slower flows, since the pressure driving the flow is smaller for the slower speed. For the pressure system, the drops led to a 1.4 mbar change in pressure for the 2.4 mm/s speed (28.3 mbar average), and a 1.1 mbar change in pressure for the 4.8 mm/s speed (47.4 mbar average). A larger pressure change for a lower average pressure produces to a larger change in flow, which agrees with the larger magnitude flow drop seen for the slower speeds. Overall, these plots demonstrate that the user should be aware of the position of the outlet line during microfluidic data acquisition, and that submersion of the outlet line is critical to eliminate rapid flow changes caused by dripping from the outlet line.

>>>>

\usepackage[letterpaper,margin=1in]{geometry}
\usepackage{authblk}
\usepackage{graphicx}

@@

An important methodological detail taken into account during all experiments was the submersion of the microfluidic outlet tubing in solution as depicted in the system schematics in Figs. 2 and 3. This strategy was adopted after visualizing dramatic changes in flow caused by the effluent dripping into the collection vial when the outlet line was elevated above the solution. The release of drops from the end of the tubing causes a large decrease in pressure, resulting in a flow decrease. This was observed with both flow control systems as shown in Fig.~\ref{fig:supplement_outlet}, indicating that this is an important experimental consideration for any microfluidic setup. These trials spanned ten minutes, with approximately five minutes of the effluent dripping into the collection vial before manually submerging the outlet line in the collected solution.

% Figure S1 - Outlet Line Submersion
\begin{figure}
    \includegraphics[width=\textwidth]{FigureS1.pdf}
    \caption {
        Impact of submerging the microfluidic device outlet tubing for both flow control systems at 2.4 and 4.8 mm/s as measured with the inline flow sensor (blue) and single-exposure LSCI (red). These trials show approximately five minutes of effluent dripping into the collection vial, as evidenced by the periodic flow decreases, followed by five minutes of operation with the outlet tubing submerged in the collected solution. The large spike in relative flow around $t = $ 300 seconds corresponds to the outlet line being manually submerged in solution. All measurements were normalized to the average flow during the submerged segment of each trial and compared to the ideal flow (black).
    }
    \label{fig:supplement_outlet}
\end{figure}

Some general trends were observed across these trials. First, the magnitude of the flow decrease was larger for the syringe pump compared to the pressure-regulated flow system. Because the pressure-regulated system has a higher compliance than the syringe pump, high frequency perturbations are dampened more by the pressure-regulated system. Second, the frequency of the droplets was very regular within individual trials. This suggests that drops must reach a critical volume to overcome the surface tension in order to release from the tubing. Using the absolute flow rate from the inline sensor and the time between drops to estimate the mean volume, a standard deviation of \textless0.5 $\mu$L in drop volume was observed within each trial. Finally, the magnitude of the flow decrease was larger for the slower flow (2.4 mm/s) than the faster flow (4.8 mm/s) across all trials. This was likely caused by the larger pressure differential generated at slower flows, since the pressure driving the flow was lower for the slower flow rate. For example, with the pressure-regulated system, the drops led to a 1.4 mbar change at the 2.4 mm/s speed (28.3 mbar average) and a 1.1 mbar change at the 4.8 mm/s speed (47.4 mbar average). The larger relative pressure differential (4.95\% vs. 2.3\%) produced a larger change in flow, which corresponds to the higher magnitude flow decreases seen at the slower flow rate. Overall, these results demonstrate that users should be aware of the positioning of the microfluidic outlet line during experiments and that submersion of the outlet tubing is critical to eliminate periodic flow perturbations caused by dripping.
